Excuse me, do you have a moment to talk about version control?

### 2017-05-19 README.Rmd
Revisions, citations, links

Git has been re-purposed by the data science community. In addition to using it for source code, we use it to manage the motley collection of files that make up typical data analytical projects, which often consist of data, figures, reports, and, yes, source code. Even those who identify more as statistician than data scientist generally have a similar mix of files that are the artifacts of a project.

@@

In my opinion, **for new users**, the pros of Git only outweigh the cons when you factor in the overhead of communicating and collaborating with other people, including your future self. And who among us does not need to do that? Your life is much easier if this is baked into your workflow, as opposed to being a separate process that you dread and avoid. Communication and collaboration are the killer apps of version control. Git enforces a rigorous model of file management, but it is critical for the distribution of files across different people, computers, and time.

@@

cat("*make such a phylogeny with a set of crazy filenames like this*\n")
cat("![](https://www.onshape.com/hs-fs/hub/379003/file-2533844795-jpg/Blog_Assets/Version_Control_Graphic-2.jpg?t=1494963046029)")

cat("*or riff on the famous PhD Comic*")
cat("![](http://phdcomics.com/comics/archive/phd101212s.gif)")
```

The Git way is to track the evolution of that file, through a series of commits, each equipped with an explanatory message and a nickname. All collaborators sync regularly to a common version, acknowledging that the difficulty of merging goes up faster than the size of the difference (cite good enough). Especially important versions get a human-readable tag, to signal a meaningful milestone. Yes, there is some pain in adopting the formalism of Git, but it is worth it.

*another good image to emulate, from the Bartlett Git for Designers talk*
![](bartlett-copies-vs-commits.png)

@@

@@

  1. When you are new with Git (or, frankly, even when you're not), it's common to goof up the Git infrastructure for a project. Note that your files can absolutely be intact and safe, even while the Git tracking is a bit confused. Of course there are official Git remedies, but sometimes the easiest fix is to clone a fresh copy from GitHub, patch things up with the changes that only exist locally, and move on with your life. This workaround obviously requires the existence of a recent copy on GitHub.
  2. The highly functional web interfaces mentioned above are often the most pleasant and natural way to navigate and search your files, even though all the same information exists locally. It is a pleasure to browse through your own work, across multiple projects or files and across time, as if it's a well-designed website. You must push your work to GitHub to enjoy this.

GitHub __issues__ are another powerful feature of the platform. Recall that we are repurposing Git, a tool designed to facilitate software development. The issues for a project are its bug tracker. For projects that are not pure software development, we co-opt this machinery to organize our to do list more generally. The basic unit is an issue and you can interact with them in two ways.

@@

This is one-time or once-per-computer setup.

@@

  * Install Git. Depending on your OS, Git might already be installed. But many of us will need to install it or might choose to update to the most recent version. Some basic configuration is critical, such as setting your username and email.

  * Install a local Git client. Optional but highly recommended. A Git client is software that provides a graphical user interface for Git, which is otherwise command-line only. If you are an R user, you will find that RStudio provides a great deal of this functionality. There are some notable gaps, however, so you might still choose to install a dedicated and comprehensive Git client such as SourceTree or GitKraken. Git is a file-based system, so you can do some operations from RStudio, others from SourceTree, and others from the shell.

  * Confirm, with a practice repository, that local Git can send and receive the current version of the repository on GitHub, known as *pushing* and *pulling*, respectively. This will require authenticating yourself with GitHub from a local shell or Git client. At this point, most people elect to do a bit of extra setup to ensure that they are not repeatedly challenged for their GitHub credentials going forward.

@@

Git was built neither for the exact usage described here, nor for broad usability. You will undoubtedly notice this, so it can be helpful to know this in advance. Happily there are many helpful tools that mediate your interactions with Git. GitHub itself is a fine example. In addition to pointing out tools that soften Git's sharpest edges, I recommend specific habits and attitudes that reduce frustration.

@@

  * I repeat: consider using a graphical front-end for Git, a.k.a. a "Git client", versus restricting yourself to the command line interface.
  * Establish confidence in the basics (e.g. make a change, commit it, push it) before wading into more advanced usage (e.g. branching).
  * Commit yourself to Git usage on a project that will provide sustained practice over several months. Usage in a course is great, because it provides a relentless stream of small deadlines.
  * Realize that no one is giving out Git style points. It's ok to "power-cycle", i.e. re-initialize the Git repository, to get unstuck.

@@

  * Dedicate a local directory (a.k.a "folder") to it.
  * Make it an RStudio Project. Optional but recommended; obviously only applies to projects involving R and users of RStudio.
  * Make it a Git repository.

This is once-per-project setup and can happen at project inception or at any later point. Chances are your project already lives in a dedicated directory. Making this directory an RStudio Project and Git repository boils down to allowing those applications to leave notes for themselves in hidden files or directories. The project is still a regular directory on your computer, that you can locate, name, move, and generally interact with as you wish. You don't have to handle it with special gloves!

@@

  * Go about your usual business, e.g. writing R scripts or authoring reports in LaTeX or R Markdown. But instead of only *saving* individual files, periodically you make a **commit**, which takes a snapshot of all the files in the entire project.
    - Have you ever versioned a file [by adding your initials or the date](http://www.phdcomics.com/comics/archive.php?comicid=1531)? That is effectively a **commit**, albeit only for a single file: it is a version that is significant to you and that you might want to inspect or revert to later.
  * Push commits to GitHub periodically.
    - This is like sharing a document with colleagues on DropBox or sending it out as an email attachment. By pushing to GitHub, you make your work and all your accumulated progress accessible to others.

This is a moderate change to your normal, daily workflow. It feels weird at first, but quickly becomes second nature. In [STAT 545](http://stat545.com) students are required to submit all coursework via GitHub, starting in week one. Most have never seen Git before and do not identify as "programmers". It is a major topic in class and office hours for the first two weeks. Then we practically never discuss it again.

*I am not saying anything about pulling and collaboration here, which seems a bit extreme, even if I can't say much*

@@

Recall that a repository -- or "repo" -- is just a directory of files that Git is going to manage holistically. A commit functions like a snapshot of all the files in the repo, at a specific moment. Under the hood, that is not exactly how Git implements things. Mental models don't have to be accurate in order to be useful, but in this case there's some value in aligning the two.

@@

![](bartlett-commit-history.png)

@@

Every time you make a commit you must also write a short *commit message*. Ideally, this conveys the *motivation* for the change. Remember, the diff will show the content. When you revisit a project after a break or need to digest recent changes made by a colleague, looking at the *history*, by reading commit messages and skimming through diffs, is an extremely efficient way to get up to speed.

Every commit needs some sort of nickname, so you can identify it. Git does this automatically, assigning each commit what is called a *SHA*, a seemingly random string of 40 letters and numbers (it is not, in fact, random but is a SHA-1 checksum hash of the commit). Though you will be exposed to these, you don't have to handle them directly very often and, when you do, usually the first 7 characters suffice. You can also designate certain snapshots as special with a *tag*, which is a name of your choosing. In a software project, it is typical to tag a release with its version, e.g., "v1.0.3". For a manuscript or analytical project, you might tag the version submitted to a journal or transmitted to external collaborators.

@@

This may seem unrelated to Git, GitHub, and R, but we need to talk about Markdown. Markdown is a markup language, like HTML and LaTeX, but designed to be as lightweight as possible. The goal is still to separate form and content, but also to prioritize human-readability, even at the cost of fancy features. Markdown is in wide use on sites like Wikipedia, WordPress, StackOverflow, and, yes, GitHub. These sites use Markdown because it allows a diverse population of site users to author decent-looking web content, with hyperlinks and some formatting. Do not build this up into some heroic, LaTeX-level learning task, for it is not. If you can write an email, you can write Markdown.

@@

Markdown is extra special for R users because of R Markdown, which is just Markdown that includes chunks of R code. Again, do not regard R Markdown as   something you must clear your schedule to learn. If you can write email and a bit of R code, you can write R Markdown. The rmarkdown package converts R Markdown (`.Rmd` files) to Markdown (`.md` files), running the code and inserting the results, including figures, into the document. This is powered by another package, knitr, under the hood. This process is made especially easy in RStudio, but is by no means limited to users of that application.

@@

These R-derived Markdown files, if committed and pushed, then enjoy the usual privileged treatment on GitHub already described above. Once an `.Rmd` file has been rendered to `.md`, anyone viewing it on GitHub can read the prose, study the R code, **and view the results of running that code**, including figures. It is the best of all worlds, because the code is revealed and, by definition, is the code that produced the results. And yet a reader can gaze upon the product in a web browser, without needing to download the code, install all necessary dependencies, and run it.

@@

You do not, in fact, even need to work in R Markdown to exploit this. It works with plain R scripts as well. You can use exactly the same machinery to prepare a rendered version of an R script, i.e. to go from `.R` to `.md`. Again, RStudio makes this especially easy, but this is not limited to RStudio. Once the Markdown file is pushed to GitHub, it is as if the reader has run your code or is able to look over your shoulder at your R session. This provides an lightweight system for exposing work-in-progress to collaborators, without slowing down to create separate reports. Comment lines that begin with `#'` are elevated to top-level prose, providing a way to make the document more welcoming for a reader. Once there are many prose comments, you might decide to switch from `.R` to `.Rmd`, have proper top-level prose, and move the code down into chunks.

@@

Let's zoom back out again and consider R Markdown and the rmarkdown package more generally. I want to point out that R Markdown can be rendered to many more formats than Markdown! I have emphasized the production of Markdown here because it is extremely useful in GitHub-hosted projects and that seems to be underappreciated. But the rmarkdown package can convert `.Rmd` to a wide array of output formats, including HTML, PDF, and Word (`.docx`). The Markdown must be created regardless. Sometimes it is all you need, such as on GitHub. In other contexts, it is just a necessary intermediate and may even be discarded.

@@

  * R markdown `.Rmd` --> markdown `.md`
  * R markdown `.Rmd` --> markdown `.md` --> `.html` or `.pdf` or `.docx`
  * R script `.R` --> results as `.csv` or `.rds` and figures `.png`
  * LaTeX `.tex` --> WHAT GOES HERE NOW? --> `.pdf`

The files at the far left are clearly source files. In the case of an R script, this is literally true, but it's morally true for R markdown and LaTeX files too. These are files that you directly create and edit "by hand".

The files at the far right are clearly derived and are often described as "targets". These files are programmatically generated from source files (and possibly other inputs). These files are the product and they have external value, often for communicating ideas and results.

@@

A critical issue for workflow happiness is figuring out how to manage the production and storage of source, intermediates, and targets with respect to Git. You can direct Git to ignore specific files or certain types of files, such as autosaves created by your editor. This reduces clutter in your project: Git will not pester you to commit changes to these files and they will not appear in the associated GitHub repository. A file that Git does not ignore is said to be *tracked*.

@@

@@

*Will it make collaboration harder?* *talk about the scenario Nick brought up re: binary files like PDF being a common source of merge conflict <https://github.com/dsscollection/git-github-for-stats/issues/6>*

@@

Collaboration is probably the most compelling reason to manage a project with Git and GitHub. I have a broad definition of collaboration here, that includes hands-on participation by multiple people as well as an asymmetric model, in which some people are active makers and others only read or review.

@@

Managing a project via Git/GitHub is much more like the Google Doc scenario, but also offers some of the attractive features of "edit, save, attach". With Git/GitHub, collaborators can work offline and there can be independent lines of development. The real power comes from regular and structured reconciliation of all the files in the project via Git/GitHub. It is definitely more complicated than collaborating on a Google Doc, but also more powerful.

@@

The best way to deal with merge conflicts is to avoid them altogether. This is another reason for all parties to commit, pull, and push often. Small changes, being integrated frequently, in non-binary files, are the easiest for Git to automatically merge for you. The difficulty of merging (by Git or by you) is proportional to the evolutionary distance between two lines of work. The presence of frequently-changing binary files also increases the burden. So make lots of small commits, sync regularly with GitHub, and only track binary files with good reason.

@@

![](bartlett-merge-commit.png)

![](bartlett-pull.png)

@@

Transparency about [@greenwade93] process and product is increasingly important in science. The SOMETHING for reproducibility is well accepted. A more underappreciated benefit is democratization of our field, as this affords a much broader audience a clear view of how scientists and programmers work.

@@

Does creation and presence in GitHub repo make the project more readable and navigable? If so DO IT.

need to add a ton of links, citations

make some bespoke examples or diagrams?

Doing = documenting = sharing

@@

[RStudio IDE](https://www.rstudio.com/products/rstudio-desktop/).

>>>>

Git has been re-purposed by the data science community [@Ram2013;
@git-for-humans]. In addition to using it for source code, we use it to manage 
the motley collection of files that make up typical data analytical projects, 
which often consist of data, figures, reports, and, yes, source code. Even those
who identify more as statistician than data scientist generally have a similar 
mix of files that are the artifacts of a project.

@@

In my opinion, __for new users__, the pros of Git only outweigh the cons when 
you factor in the overhead of communicating and collaborating with other people,
including your future self. And who among us does not need to do that? Your life
is much easier if this is baked into your workflow, as opposed to being a 
separate process that you dread and avoid. Communication and collaboration are 
the killer apps of version control. Git enforces a rigorous model of file 
management, but it is critical for the distribution of files across different 
people, computers, and time.

@@

cat("\n\n*make such a phylogeny with a set of crazy filenames like this*\n") 
cat("![](https://www.onshape.com/hs-fs/hub/379003/file-2533844795-jpg/Blog_Assets/Version_Control_Graphic-2.jpg?t=1494963046029)")

cat("\n\n*or riff on the famous PhD Comic*") 
cat("![](http://phdcomics.com/comics/archive/phd101212s.gif)")
```

The Git way is to track the evolution of that file, through a series of commits,
each equipped with an explanatory message and a nickname. All collaborators sync
regularly to a common version, acknowledging that the difficulty of merging goes
up faster than the size of the difference [@good-enough]. Especially important 
versions get a human-readable tag, to signal a meaningful milestone. Yes, there 
is some pain in adopting the formalism of Git, but it is worth it.

*make a figure related to the above (reference those file names?) but 
juxtaposing different workflows, inspiration from the Bartlett Git for Designers
talk*

```{r echo = FALSE, out.width="100%"}
knitr::include_graphics("bartlett-copies-vs-commits.png")
```

@@

*possible images include a modified version of this atlassian image (no arrows 
between people, yes arrow from people to github), somehow connected to a 
screenshot of a GitHub repo*

```{r echo = FALSE, out.width="100%"}
knitr::include_graphics("atlassian-git-tutorials-syncing.png")
```

@@

1. When you are new with Git (or, frankly, even when you're not), it's common to
goof up the Git infrastructure for a project. Note that your files can be intact
and safe, even while the Git tracking is a bit confused. Of course there are 
official Git remedies, but sometimes the easiest fix is to clone a fresh copy 
from GitHub, patch things up with the changes that only exist locally, and move 
on with your life. This workaround obviously requires the existence of a recent 
copy on GitHub.

2. The highly functional web interfaces mentioned above are often the most 
pleasant and natural way to navigate and search your files, even though all the 
same information exists locally. It is a pleasure to browse through your own 
work, across multiple projects or files and across time, as if it's a 
well-designed website. You must push your work to GitHub to enjoy this.

__GitHub issues__ are another powerful feature of the platform. Recall that we 
are repurposing Git, a tool designed to facilitate software development. The 
issues for a project are its bug tracker. For projects that are not pure 
software development, we co-opt this machinery to organize our to do list more 
generally. The basic unit is an issue and you can interact with one in two ways.

@@

Here is the initial setup process, which you do once or, for some steps, once per computer.

@@

* Install Git. Depending on your OS, Git might already be installed. But many of
us may need to install it or will choose to update to a more recent version.
Some basic configuration is critical, such as setting your username and email.

* Install a local Git client *(optional but highly recommended*). A Git client 
is software that provides a graphical user interface for Git, which is otherwise
command-line only. If you are an R user, you will find that 
[RStudio](https://www.rstudio.com/products/rstudio-desktop/) provides a great 
deal of this functionality. There are some notable gaps, however, so you might 
still choose to install a dedicated and comprehensive Git client such as 
[SourceTree](https://www.sourcetreeapp.com) or
[GitKraken](https://www.gitkraken.com). Git is a file-based system, so you can
do some operations from RStudio, others from SourceTree, and others from the
shell.

* Confirm, with a practice repository, that local Git can send and receive the 
current version of the repository on GitHub, known as __pushing__ and __pulling__, respectively. This will require authenticating yourself with GitHub from a local shell or Git client. At this point, most people elect to do a bit of extra setup to ensure that they are not repeatedly challenged for their GitHub credentials going forward.

@@

Git was built neither for the exact usage described here, nor for broad 
usability. You will undoubtedly notice this, so it can be helpful to know this 
in advance. Happily there are many helpful tools that mediate your interactions 
with Git. GitHub itself is a fine example, as is RStudio. In addition to pointing out tools that soften Git's sharpest edges, I recommend specific habits and attitudes that reduce frustration.

@@

  * I repeat: consider using a graphical front-end for Git, a.k.a. a Git client, versus restricting yourself to the command line interface.
  * Establish confidence in the basics (e.g. make a change, commit it, push it) before wading into more advanced usage (e.g. branching).
  * Commit yourself to Git usage on a project that will provide sustained practice over several months. Usage in a course is great, because it provides a relentless stream of small deadlines.
  * Realize that no one is giving out Git style points. It's ok to "power-cycle", i.e. re-initialize the Git repository, to get unstuck.

@@

  * Dedicate a local directory or folder to it.
  * Make it an RStudio Project. *Optional but recommended; obviously only applies to projects involving R and users of RStudio.*
  * Make it a Git repository.

This setup happens once per project and can happen at project inception or at any later point. Chances are your project already lives in a dedicated directory. Making this directory an RStudio Project and Git repository boils down to allowing those applications to leave notes for themselves in hidden files or directories. The project is still a regular directory on your computer, that you can locate, name, move, and generally interact with as you wish. You don't have to handle it with special gloves!

@@

  * Go about your usual business, writing R scripts or authoring reports in LaTeX or R Markdown. But instead of only *saving* individual files, periodically you make a __commit__, which takes a snapshot of all the files in the entire project.
    - If you have ever versioned a file [by adding your initials or the date](http://www.phdcomics.com/comics/archive.php?comicid=1531), you have effectively made a commit, albeit only for a single file. It is a version that is significant to you and that you might want to inspect or revert to later.
  * Push commits to GitHub periodically.
    - This is like sharing a document with colleagues on DropBox or sending it out as an email attachment. By pushing to GitHub, you make your work and all your accumulated progress accessible to others.

This is a moderate change to your normal, daily workflow. It feels weird at first, but quickly becomes second nature. In [STAT 545](http://stat545.com) students are required to submit all coursework via GitHub, starting in week one. Most have never seen Git before and do not identify as programmers. It is a major topic in class and office hours for the first two weeks. Then we practically never discuss it again.

@@

Recall that a repository or repo is just a directory of files that Git will manage holistically. A commit functions like a snapshot of all the files in the repo, at a specific moment. Under the hood, that is not exactly how Git implements things. Mental models don't have to be accurate in order to be useful, but in this case there's some value in aligning the two.

@@

```{r echo = FALSE, out.width="100%"}
knitr::include_graphics("bartlett-commit-history.png")
```

@@

Every time you make a commit you must also write a short __commit message__. Ideally, this conveys the motivation for the change. Remember, the diff will show the content. When you revisit a project after a break or need to digest recent changes made by a colleague, looking at the __history__, by reading commit messages and skimming through diffs, is an extremely efficient way to get up to speed.

Every commit needs some sort of nickname, so you can identify it. Git does this automatically, assigning each commit what is called a SHA, a seemingly random string of 40 letters and numbers (it is not, in fact, random but is a SHA-1 checksum hash of the commit). Though you will be exposed to these, you don't have to handle them directly very often and, when you do, usually the first 7 characters suffice. You can also designate certain snapshots as special with a __tag__, which is a name of your choosing. In a software project, it is typical to tag a release with its version, e.g., "v1.0.3". For a manuscript or analytical project, you might tag the version submitted to a journal or transmitted to external collaborators.

@@

This may seem unrelated to Git, GitHub, and R, but we need to talk about [Markdown](https://daringfireball.net/projects/markdown/syntax). Markdown is a markup language, like HTML and LaTeX, but designed to be as lightweight as possible. The goal is still to separate form and content, but also to prioritize human-readability, even at the cost of fancy features. Markdown is in wide use on sites like [WordPress](https://en.support.wordpress.com/markdown/), [StackOverflow](https://stackoverflow.com/editing-help), and, yes, [GitHub](https://help.github.com/categories/writing-on-github/). These sites use Markdown because it allows a diverse population of site users to author decent-looking web content, with hyperlinks and some formatting. Do not build this up into some heroic, LaTeX-level learning task, for it is not. If you can write an email, you can write Markdown.

@@

Markdown is extra special for R users because of [R Markdown](http://rmarkdown.rstudio.com), which is just Markdown that includes chunks of R code. Again, do not regard R Markdown as something you must clear your schedule to learn. If you can write email and a bit of R code, you can write R Markdown. The [rmarkdown package](https://CRAN.R-project.org/package=rmarkdown) [@rmd-pkg] converts R Markdown (`.Rmd` files) to Markdown (`.md` files), running the code and inserting the results, including figures, into the document. This is powered by another package, [knitr](https://CRAN.R-project.org/package=knitr) [@knitr-pkg; @knitr-book], under the hood. This process is made especially easy in RStudio, but is by no means limited to users of that application.

@@

These R-derived Markdown files, if committed and pushed, then enjoy the usual privileged treatment on GitHub already described above. Once an `.Rmd` file has been rendered to `.md`, anyone viewing it on GitHub can read the prose, study the R code, __and view the results of running that code__, including figures. It is the best of all worlds, because the code is revealed and, by definition, is the code that produced the results. And yet a reader can gaze upon the product in a web browser, without needing to download the code, install all necessary dependencies, and run it.

@@

You do not, in fact, even need to work in R Markdown to exploit this. It [works with plain R scripts](http://rmarkdown.rstudio.com/articles_report_from_r_script.html) as well. You can use exactly the same machinery to prepare a rendered version of an R script, i.e. to go from `.R` to `.md`. Again, RStudio makes this especially easy, but this is not limited to RStudio. Once the Markdown file is pushed to GitHub, it is as if the reader has run your code or is able to look over your shoulder at your R session. This provides an lightweight system for exposing work-in-progress to collaborators, without slowing down to create separate reports. Comment lines that begin with `#'` are elevated to top-level prose, providing a way to make the document more welcoming for a reader. Once there are many prose comments, you might decide to switch from `.R` to `.Rmd`, have proper top-level prose, and move the code down into chunks.

@@

Let's zoom back out again and consider R Markdown and the rmarkdown package more generally. I want to point out that R Markdown can be rendered to many more formats than Markdown! I have emphasized the production of Markdown here because it is extremely useful in GitHub-hosted projects and that seems to be underappreciated. But the rmarkdown package can convert `.Rmd` to a wide array of [output formats](http://rmarkdown.rstudio.com/lesson-9.html), including HTML, PDF, and Word (`.docx`). The Markdown must be created regardless. Sometimes it is all you need, such as on GitHub. In other contexts, it is just a necessary intermediate and may even be discarded.

@@

  * R markdown `.Rmd` --> markdown `.md`
  * R markdown `.Rmd` --> markdown `.md` --> `.html` or `.pdf` or `.docx`
  * R script `.R` --> results as `.csv` or `.rds` and figures `.png`
  * LaTeX `.tex` --> `.aux`, `.bbl`, etc. --> `.pdf`

The files at the far left are clearly source files. In the case of an R script, this is literally true, but it's morally true for R markdown and LaTeX files too. These are files that you directly create and edit by hand.

The files at the far right are clearly derived and are often described as __targets__. These files are programmatically generated from source files (and possibly other inputs). These files are the product and they have external value, often for communicating ideas and results.

@@

A critical issue for workflow happiness is figuring out how to manage the production and storage of source, intermediates, and targets with respect to Git. You can direct Git to ignore specific files or certain types of files, such as autosaves created by your editor. This reduces clutter in your project: Git will not pester you to commit changes to these files and they will not appear in the associated GitHub repository. A file that Git does not ignore is said to be __tracked__.

@@

*I suspect this is a good place to be less wordy? I am so frustrated with people doing stupid things here, though!*

@@

*Will it make collaboration harder?* Prose to be written: Talk about the scenario Nick brought up re: binary files like PDF being a common source of merge conflict <https://github.com/dsscollection/git-github-for-stats/issues/6>. I have a solution! Stop using PDF as your default output format! Make Markdown the output format during development. Your Git/GitHub problems just went away. You're welcome.

@@

Collaboration is probably the most compelling reason to manage a project with Git and GitHub. I have a broad definition of collaboration. It includes hands-on participation by multiple people as well as an asymmetric model, in which some people are active makers and others only read or review.

@@

Managing a project via Git/GitHub is much more like the Google Doc scenario, but also offers some of the attractive features of "edit, save, attach". With Git/GitHub, collaborators can work offline and there can be independent lines of development. The real power comes from regular and structured reconciliation of all versions of all the files in the project via Git/GitHub. It is definitely more complicated than collaborating on a Google Doc, but also more powerful.

@@

The best way to deal with merge conflicts is to avoid them altogether. This is another reason for all parties to commit, pull, and push often. Small changes, integrated frequently, in non-binary files, are the easiest for Git to automatically merge for you. The difficulty of merging (by Git or by you) is proportional to the evolutionary distance between two lines of work. The presence of frequently-changing binary files also increases the burden. So make lots of small commits, sync regularly with GitHub, and only track binary files with good reason.

@@

```{r echo = FALSE, out.width="100%"}
knitr::include_graphics("bartlett-merge-commit.png")
```

```{r echo = FALSE, out.width="100%"}
knitr::include_graphics("bartlett-pull.png")
```

@@

Transparency about process and product is increasingly important in science. The SOMETHING for reproducibility is well accepted. A more underappreciated benefit is democratization of our field, as this affords a much broader audience a clear view of how scientists and programmers work.

@@

Consider links to GitHub repos that exemplify certain points and are very unlikely to disappear any time soon.n

@@
